---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Riccardo Fusaroli"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loading libraries
library(pacman)
p_load(tidyverse, lme4, groupdata2, caret, tidymodels)

# Loading data
data_dan <- read.csv("data_dan.csv", header = T, sep = ",")
data_chi <- read.csv("data_chi.csv", header = T, sep = ",")
data_jap <- read.csv("data_jap.csv", header = T, sep = ",")

# Assign languages
data_dan$language <- "Danish"
data_chi$language <- "Chinese"
data_jap$language <- "Japanese"

# Create single dataframe and clean environment
data <- rbind(data_dan, data_chi, data_jap)
rm(data_dan, data_chi, data_jap)

# Remove columns from train set
data <- data %>%
  select(-Filename, -uPairID)

# Set seed for randomization
set.seed(5)

```

## Assignment 3 - Part 2 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia. We then looked at whether we could replicate results from the previous literature.
We now want to know whether we can automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.
Again, remember that the dataset containst 7 studies and 3 languages. Feel free to only include Danish (Study 1-4) if you feel that adds too much complexity.

Issues to be discussed your report:
- Should you run the analysis on all languages/studies at the same time? 
- Choose your best acoustic feature from part 1. How well can you diagnose schizophrenia just using it?
- Identify the best combination of acoustic features to diagnose schizophrenia using logistic regression.
- Discuss the "classification" process: which methods are you using? Which confounds should you be aware of? What are the strength and limitation of the analysis?
- Bonus question: Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them. 
- Bonus Bonus question: It is possible combine the output of multiple  classification models to improve classification accuracy. For inspiration see,
https://machinelearningmastery.com/machine-learning-ensembles-with-r/
 The interested reader might also want to look up 'The BigChaos Solution to the Netflix Grand Prize'

## Learning objectives
- Learn the basics of classification in a machine learning framework
- Design, fit and report logistic regressions
- Apply feature selection techniques

### Let's start

We first want to build a logistic regression to see whether you can diagnose schizophrenia from your best acoustic feature. Let's use the full dataset and calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve). You need to think carefully as to how we should (or not) use study and subject ID.
```{r}
### Preparing Testing and Training sets ###
# Create test sets and training sets from each dataframe
test_train_list <- partition(data, p = 0.2, cat_col = c("Diagnosis", "language"), id_col = "uID", list_out = T) # splits data according to ID's and ensures same distribution of Diagnosis and language
# Create the testing and training set
test = test_train_list[[1]]
train = test_train_list[[2]]
rm(test_train_list) # Clean environment



### Logistic Regression from best feature ###
# Create and fit the model on the training set
log_reg <- logistic_reg() %>%
  set_mode("classification") %>% 
  set_engine("glm") %>%
  fit(Diagnosis ~ iqr + (1|uID), data = train)

# Create a dataframe with true versus predicted values as well as probability values
log_reg_results <- test %>%
  select(Diagnosis) %>% 
    mutate(
      pred_class = predict(log_reg, new_data = test) %>% 
        pull(.pred_class),
      prob  = predict(log_reg, new_data = test, type = "prob") %>% 
        pull(.pred_SCZ))


# Making the ROC curve
roc_curve(data = log_reg_results, truth = Diagnosis, prob) %>% 
  autoplot()


# Get performance measures
confmat <- confusionMatrix(log_reg_results$Diagnosis, log_reg_results$pred_class)
confmat

```


Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures. Alternatively, the groupdata2 and cvms package created by Ludvig are an easy solution.

N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one dataset, so to calculate overall performance.
N.N.N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?
N.N.N.N.B. A more advanced solution could rely on the tidymodels set of packages (warning: Time-consuming to learn as the documentation is sparse, but totally worth it)



```{r}

# Create folds for cross validation
# Initialize fold column
data$fold <- 0

# Loop for assigning folds based on unique ID's and language
for (i in c("Danish", "Chinese", "Japanese")){
  data$fold[data$Diagnosis == "CON" & data$language == i] <-
    createFolds(unique(data$uID[data$Diagnosis == "CON" & data$language == i]), k = 5, list = FALSE)

  data$fold[data$Diagnosis == "SCZ" & data$language == i] <-
    createFolds(unique(data$uID[data$Diagnosis == "SCZ" & data$language == i]), k = 5, list = FALSE)
}

# Create dataframe for storing true and predicted values
cv_results <- table("Diagnosis", "fold", "pred_class", "prob")

# Crossvalidate and save predictions using for loop
for (i in 1:max(data$fold)){
  # Create test sets and training sets based on current iteration
  cv_test <- data[data$fold == i,]
  cv_train <- data[data$fold != i,]
  
  #logistic regression on train data
  cv_log_reg <- 
    logistic_reg() %>%
    set_mode("classification") %>%
    set_engine("glm") %>%
      fit(Diagnosis ~ iqr + (1|uID), data = cv_train)
  
  #get multiple at once
  fold_results <- 
    cv_test %>% 
    select(Diagnosis, fold) %>% 
    mutate(
      pred_class = predict(cv_log_reg, new_data = cv_test) %>% 
        pull(.pred_class),
      prob  = predict(cv_log_reg, new_data = cv_test, type = "prob") %>% 
        pull(.pred_SCZ)
    )
  
  # bind results to new dataframe
  cv_results <- rbind(cv_results, fold_results)
  
  # Clean environemnt
  rm(cv_train, cv_test, cv_log_reg, fold_results)
}

# remove first row of results dataframe
cv_results <- cv_results[-1,]

# Create ROC curve
roc_curve(data = cv_results, truth = Diagnosis, prob) %>% 
  autoplot()

# Get performance measures
confmat <- confusionMatrix(cv_results$Diagnosis, cv_results$pred_class)
confmat

```



Identify best features for logistic regression
``` {r}

#logistic regression on train data
log_reg <- logistic_reg() %>%
  set_mode("classification") %>% 
  set_engine("glm") %>%
  fit(Diagnosis ~ iqr + speech_rate + dur_pause:language + (1|uID), data = train)


# Create a dataframe with true versus predicted values as well as probability valuestest_results <- 
log_reg_results <- test %>%
  select(Diagnosis) %>% 
  mutate(
    pred_class = predict(log_reg, new_data = test) %>% 
      pull(.pred_class),
    prob  = predict(log_reg, new_data = test, type = "prob") %>% 
      pull(.pred_SCZ),
  )


# Making the ROC curve
roc_curve(data = log_reg_results, truth = Diagnosis, prob) %>% 
  autoplot()


# Get performance measures
confmat <- confusionMatrix(log_reg_results$Diagnosis, log_reg_results$pred_class)
confmat


```